Trump vs. Harris: A 2024 US Presidential Elections Modelling Forecast

### 2024-11-04 paper/paper.qmd
ran through styler

```{r}
#| label: fig-percentage
#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
#| echo: false
#| eval: true
#| warning: false
#| 
#| 
#| message: false
ggplot(cleaned_data, aes(x = Percentage)) +
  geom_histogram(bins = 30,
                 fill = "orange",
                 color = "black",
                 alpha = 0.7) +
  labs(title = "Distribution of Percentage", 
       x = "Percentage",
       y = "Frequency") +
  theme_minimal()
```

@@

harris_trump_selection <- cleaned_data %>%
  filter(CandidateName %in% c("Donald Trump", 
                              "Kamala Harris"))

@@

ggplot(combined_data, aes(x = CandidateName, 
                          y = Percentage, 
                          fill = CandidateName)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
  labs(title = "Percentage by Candidate", 
       x = "Candidate", 
       y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("grey", "salmon", "blue")) +
  facet_wrap(~CandidateName, scales = "free_x") +
  theme_minimal()
```

@@

ggplot(Select_candidates_cleaned_data, aes(x = PollScore)) +
  geom_histogram(aes(y = ..density..),
                 bins = 30,
                 fill = "blue",
                 color = "black",
                 alpha = 0.7) +
    geom_density(color = "black", fill = "salmon", alpha = 0.6) +
  labs(title = "Distribution of Poll Scores by Candidate",
       x = "Poll Score", 
       y = "Frequency") +
  theme_minimal() +
  facet_wrap(~ CandidateName)
```

@@

ggplot(cleaned_data, aes(x = SampleSize)) +
  geom_histogram(bins = 30,
                 fill = "green",
                 color = "black",
                 alpha = 0.7) +
  labs(title = "Distribution of Sample Size",
       x = "Sample Size", 
       y = "Frequency")+
  theme_minimal()
```

@@

ggplot(cleaned_data, aes(x = PollRecency)) +
  geom_histogram(bins = 30,
                 fill = "pink",
                 color = "black",
                 alpha = 0.7) +
  labs(title = "Distribution of Poll Recency",
       x = "PollRecency", 
       y = "Frequency")+
  theme_minimal()
```

@@

ggplot(cleaned_data, aes(x = NumericGrade)) +
  geom_histogram(bins = 30,
                 fill = "purple",
                 color = "black",
                 alpha = 0.7) +
  labs(title = "Distribution of Numeric Grade",
       x = "Numeric Grade", 
       y = "Frequency")+
  theme_minimal()
```

@@

rmse_results %>%
  kable(caption = "RMSE Results for Candidate Models",
        col.names = c("Model", "Root Mean Squared Error (RMSE)"),
        digits = 2) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```

@@

cleaned_data <- cleaned_data %>%
  mutate(PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
         PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data))

@@

results %>%
  select(NationalTrumpAverage, NormalizedTrumpPercentage, NationalHarrisAverage, NormalizedHarrisPercentage) %>%
  kable(
    caption = "Predicted National Popular Vote for Donald Trump and Kamala Harris",
    col.names = c("Trump Average Predicted %", "Trump Normalized %", "Harris Average Predicted %", "Harris Normalized %"),
    digits = 2,
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```

@@

prediction_data <- cleaned_data %>%
  group_by(State) %>%
  summarise(
    PollScore = mean(PollScore, na.rm = TRUE),
    SampleSize = mean(SampleSize, na.rm = TRUE),
    NumericGrade = mean(NumericGrade, na.rm = TRUE),
    PollRecency = mean(PollRecency, na.rm = TRUE),
    .groups = 'drop'
  )

@@

electoral_votes <- tibble(
  State = c("Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana", 
            "Ohio", "Arizona", "Maine", "Oklahoma", "Arkansas", 
            "Maryland", "Oregon", "California", "Massachusetts", "Pennsylvania", 
            "Colorado", "Michigan", "Rhode Island", "Connecticut", "Minnesota", 
            "South Carolina", "Delaware", "Mississippi", "South Dakota", 
            "District of Columbia", "Missouri", "Tennessee", "Florida", 
            "Montana", "Texas", "Georgia", "Nebraska", "Utah", 
            "Hawaii", "Nevada", "Vermont", "Idaho", "New Hampshire", 
            "Virginia", "Illinois", "New Jersey", "Washington", 
            "Indiana", "New Mexico", "West Virginia", "Iowa", 
            "New York", "Wisconsin", "Kansas", "North Carolina", 
            "Wyoming"),
  Votes = c(9, 8, 3, 3, 8, 
            17, 11, 4, 7, 6, 
            10, 8, 54, 11, 19, 
            10, 15, 4, 7, 10, 
            9, 3, 6, 3, 
            3, 10, 11, 30, 
            4, 40, 16, 5, 6, 
            4, 6, 3, 4, 4, 
            13, 19, 14, 12, 
            11, 5, 4, 6, 
            28, 10, 6, 16, 
            3)
)

@@

state_winners %>%
  select(State, TrumpPredicted, HarrisPredicted, Winner) %>%
  kable(
    caption = "Predicted Percentage of Electoral Votes Per State (Margin ≤ 4%)",
    col.names = c("State", "Trump Predicted %", "Harris Predicted %", "Winner"),
    digits = 2,
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```

@@

map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Winner), color = "white") +
  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue")) +
  labs(title = "2024 US Presidential Election Predictions by State",
       fill = "Predicted Winner") +
  theme_void() +
  theme(legend.position = "right")

@@

summary_stats %>%
  kable(
    caption = "Summary Statistics of Model Results",
    col.names = c("Average Poll Score", "Average Sample Size", "Average Percentage", "Total Polls"),
    digits = 3,
    booktabs= TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

>>>>

```{r}
#| label: fig-percentage
#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
#| echo: false
#| eval: true
#| warning: false
#|
#|
#| message: false
ggplot(cleaned_data, aes(x = Percentage)) +
  geom_histogram(
    bins = 30,
    fill = "orange",
    color = "black",
    alpha = 0.7
  ) +
  labs(
    title = "Distribution of Percentage",
    x = "Percentage",
    y = "Frequency"
  ) +
  theme_minimal()
```

@@

harris_trump_selection <- cleaned_data %>%
  filter(CandidateName %in% c(
    "Donald Trump",
    "Kamala Harris"
  ))

@@

ggplot(combined_data, aes(
  x = CandidateName,
  y = Percentage,
  fill = CandidateName
)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
  labs(
    title = "Percentage by Candidate",
    x = "Candidate",
    y = "Percentage"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("grey", "salmon", "blue")) +
  facet_wrap(~CandidateName, scales = "free_x") +
  theme_minimal()
```

@@

ggplot(Select_candidates_cleaned_data, aes(x = PollScore)) +
  geom_histogram(aes(y = ..density..),
    bins = 30,
    fill = "blue",
    color = "black",
    alpha = 0.7
  ) +
  geom_density(color = "black", fill = "salmon", alpha = 0.6) +
  labs(
    title = "Distribution of Poll Scores by Candidate",
    x = "Poll Score",
    y = "Frequency"
  ) +
  theme_minimal() +
  facet_wrap(~CandidateName)
```

@@

ggplot(cleaned_data, aes(x = SampleSize)) +
  geom_histogram(
    bins = 30,
    fill = "green",
    color = "black",
    alpha = 0.7
  ) +
  labs(
    title = "Distribution of Sample Size",
    x = "Sample Size",
    y = "Frequency"
  ) +
  theme_minimal()
```

@@

ggplot(cleaned_data, aes(x = PollRecency)) +
  geom_histogram(
    bins = 30,
    fill = "pink",
    color = "black",
    alpha = 0.7
  ) +
  labs(
    title = "Distribution of Poll Recency",
    x = "PollRecency",
    y = "Frequency"
  ) +
  theme_minimal()
```

@@

ggplot(cleaned_data, aes(x = NumericGrade)) +
  geom_histogram(
    bins = 30,
    fill = "purple",
    color = "black",
    alpha = 0.7
  ) +
  labs(
    title = "Distribution of Numeric Grade",
    x = "Numeric Grade",
    y = "Frequency"
  ) +
  theme_minimal()
```

@@

rmse_results %>%
  kable(
    caption = "RMSE Results for Candidate Models",
    col.names = c("Model", "Root Mean Squared Error (RMSE)"),
    digits = 2
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

@@

cleaned_data <- cleaned_data %>%
  mutate(
    PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
    PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data)
  )

@@

results %>%
  select(NationalTrumpAverage, NormalizedTrumpPercentage, NationalHarrisAverage, NormalizedHarrisPercentage) %>%
  kable(
    caption = "Predicted National Popular Vote for Donald Trump and Kamala Harris",
    col.names = c("Trump Average Predicted %", "Trump Normalized %", "Harris Average Predicted %", "Harris Normalized %"),
    digits = 2,
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

@@

prediction_data <- cleaned_data %>%
  group_by(State) %>%
  summarise(
    PollScore = mean(PollScore, na.rm = TRUE),
    SampleSize = mean(SampleSize, na.rm = TRUE),
    NumericGrade = mean(NumericGrade, na.rm = TRUE),
    PollRecency = mean(PollRecency, na.rm = TRUE),
    .groups = "drop"
  )

@@

electoral_votes <- tibble(
  State = c(
    "Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana",
    "Ohio", "Arizona", "Maine", "Oklahoma", "Arkansas",
    "Maryland", "Oregon", "California", "Massachusetts", "Pennsylvania",
    "Colorado", "Michigan", "Rhode Island", "Connecticut", "Minnesota",
    "South Carolina", "Delaware", "Mississippi", "South Dakota",
    "District of Columbia", "Missouri", "Tennessee", "Florida",
    "Montana", "Texas", "Georgia", "Nebraska", "Utah",
    "Hawaii", "Nevada", "Vermont", "Idaho", "New Hampshire",
    "Virginia", "Illinois", "New Jersey", "Washington",
    "Indiana", "New Mexico", "West Virginia", "Iowa",
    "New York", "Wisconsin", "Kansas", "North Carolina",
    "Wyoming"
  ),
  Votes = c(
    9, 8, 3, 3, 8,
    17, 11, 4, 7, 6,
    10, 8, 54, 11, 19,
    10, 15, 4, 7, 10,
    9, 3, 6, 3,
    3, 10, 11, 30,
    4, 40, 16, 5, 6,
    4, 6, 3, 4, 4,
    13, 19, 14, 12,
    11, 5, 4, 6,
    28, 10, 6, 16,
    3
  )
)

@@

state_winners %>%
  select(State, TrumpPredicted, HarrisPredicted, Winner) %>%
  kable(
    caption = "Predicted Percentage of Electoral Votes Per State (Margin ≤ 4%)",
    col.names = c("State", "Trump Predicted %", "Harris Predicted %", "Winner"),
    digits = 2,
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```

@@

map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Winner), color = "white") +
  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue")) +
  labs(
    title = "2024 US Presidential Election Predictions by State",
    fill = "Predicted Winner"
  ) +
  theme_void() +
  theme(legend.position = "right")

@@

summary_stats %>%
  kable(
    caption = "Summary Statistics of Model Results",
    col.names = c("Average Poll Score", "Average Sample Size", "Average Percentage", "Total Polls"),
    digits = 3,
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))
```
